### 2013-08-07 Editorial.tex
Final copy
Signed-off-by: Edward Z. Yang <ezyang@mit.edu>

\documentclass{tmr}

\title{Editorial}

\author{Edward Z. Yang\email{ezyang@cs.stanford.edu}}

\begin{document}

Summer brings with it the release of three articles talking
about three different (and each arguably essential) ideas from
the Haskell landscape.  The first is an in-depth tutorial about
generalized algebraic data types, covering all of the classic
use-cases of this useful type system extension.  The second is
a meditation on the design pattern of monad transformers, and how
they can help the develop the design of a parser library.  The
final shows how even a simple API like that of a monoid can
help code achieve dramatic speed-ups.  Happy reading!

\end{document}